Computation and comparison of the stable northeastern United States marine boundary layer

### 2020-12-08 Paper/methodology.tex
More fixes from Mike Sprague
Modify equations section based on comments from Sprague

\subsubsection{Governing equations}
Both computational fluid dynamics (CFD) codes solve the incompressible form of
the Navier-Stokes equations with appropriate models for turbulence
closure. Equation~\ref{eqn:ns-les} shows the continuity and momentum equations
with all the terms necessary to model the atmospheric boundary layer
problem. Term $\mathbf{I}$ represents the pressure gradient forces as a
deviation from hydrostatic and horizontal mean gradients, term $\mathbf{II}$
represents the contribution from viscous and sub-filter scale stresses, term
$\mathbf{III}$ represents the contribution from Coriolis forces due to earth's
rotation, term $\mathbf{IV}$ represents the effects of buoyancy, and term
$\mathbf{V}$ represents source terms necessary to drive the flow to a desired
horizontal mean velocity at prescribed heights.

\begin{align}
  \frac{\partial \bar{\rho}} {\partial t} + \frac{\partial \bar{\rho} \widetilde{u_j}}{\partial x_j} & = 0, \nonumber\\
  \frac{\partial}{\partial t} \left(\bar{\rho}\, \widetilde{u}_i\right) +
  \frac{\partial}{\partial x_j} \left( \bar{\rho}\, \widetilde{u}_i \widetilde{u}_j \right) &=
  - \underbrace{\frac{\partial p'}{\partial x_j} \delta_{ij}}_\mathbf{I}
  - \underbrace{\frac{\partial \tau_{ij}}{\partial x_j}}_\mathbf{II}
  - \underbrace{2\bar{\rho}\,\epsilon_{ijk}\,\Omega_j\widetilde{u_k}}_\mathbf{III}
  + \underbrace{\left(\rho - \rho_\circ \right) g_i}_\mathbf{IV}
  + \underbrace{S^{u}_{i}}_\mathbf{V} \label{eqn:ns-les}
\end{align}

For the simulations presented in this paper, the buoyancy term ($\mathbf{IV}$)
is approximated using the Bousinessq model. The Bousinessq model ignores the
effect of density gradients on inertia while retaining its effects on buoyancy.
The density fluctuations governing the buoyancy effects are approximated as

\begin{align}
  \frac{\rho}{\rho_\circ} \approx 1 - \beta \left( \theta - \theta_\circ \right)
\end{align}
This leads to a forcing term that depends on potential temperature ($\theta$),
the reference density $\rho_\circ = \bar{\rho}$, and thermal expansion
coefficient $\beta \approx 1 / \theta_\circ$ as
\begin{align}
  \left(\rho - \rho_\circ \right) g_i = -\rho_\circ\, \beta\, g_i \left( \theta - \theta_\circ \right)
\end{align}

For ABL problems, the conservation of energy equation is usually written in
terms of potential temperature as shown in Eq.~\ref{eqn:pot-temp-les}.
\begin{align}
  \frac{\partial}{\partial t} \left(\bar{\rho}\, \widetilde{\theta}\right) +
  \frac{\partial}{\partial x_j} \left(\bar{\rho}\, \widetilde{u}_j \widetilde{\theta} \right) = - \frac{\partial}{\partial x_j} \hat{q}_j \label{eqn:pot-temp-les}
\end{align}
Under the assumption of ideal gas conditions and constant specific heat capacity
$c_p$, the gradients in potential temperature are proportional to the gradients
in absolute temperature, i.e.,
\begin{align}
   \left[ \frac{\partial T}{\partial t}, \frac{\partial T}{\partial x}, \frac{\partial T}{\partial y} \right] =
   \left( \frac{\bar{p}}{p_\circ} \right)^{\left(\frac{R}{c_p}\right)} \left[ \frac{\partial \theta}{\partial t}, \frac{\partial \theta}{\partial x}, \frac{\partial \theta}{\partial y} \right]
\end{align}
Furthermore, ignoring the pressure and viscous work terms in
Eq.~\ref{eqn:pot-temp-les}, for incompressible flows, it can be shown that
solving the enthalpy equation is equivalent to solving the potential temperature
equation. The enthalpy equation solved in Nalu-Wind for wind energy problems is
shown below
\begin{align}
  \frac{\partial}{\partial t} \left(\bar{\rho}\, \widetilde{h}\right) +
  \frac{\partial}{\partial x_j} \left(\bar{\rho}\, \widetilde{u}_j \widetilde{h} \right) = - \frac{\partial}{\partial x_j} q_j . \label{eqn:enthalpy-les}
\end{align}
It is noted here that the terms $\hat{q}_j$ (Eq.~\ref{eqn:pot-temp-les}) and
$q_j$ (Eq.~\ref{eqn:enthalpy-les}) are not equivalent and must be scaled
appropriately. Users can still provide the appropriate initial and boundary
conditions in terms of potential temperature field. Under these assumptions and
conditions, the resulting solution can then be interpreted as the variation of
potential temperature field in the computational domain.

@@

\begin{align}
  \frac{\partial}{\partial t}\left(\bar{\rho} k \right) + \frac{\partial}{\partial x_j} \bar{\rho} k \widetilde{u_j} &= \frac{\partial}{\partial x_j} \left (2 \nu_t \frac{\partial k}{\partial x_j} \right) - 2 \nu_t \widetilde{S_{ij}} \widetilde{S_{ij}} + \frac{g}{\theta_\circ} \tau_{\theta_i} - C_{\epsilon} \frac{k^{3/2}}{l} \label{eqn:ksgs-les}
\end{align}
where
\begin{align}
  C_\epsilon &= 0.93 & \nu_t &= C_\epsilon l \sqrt{k} \\
  \tau_{\theta i} &= - 2\ \frac{\nu_t}{\mathrm{Pr}_t}\ \frac{\partial \widetilde{\theta}}{\partial x_i} & \mathrm{Pr}_t &= \left( 1 + \frac{2l}{\Delta} \right)^{-1} \\
\end{align}
The choice of the length scale ($l$) depends on whether the stratification is positive
or negative, based on the recommendation by Moeng~\cite{Moeng1984}, and is given
by
\begin{align}
  l =
  \begin{cases}
    \Delta = \sqrt[3]{\Delta_x \Delta_y \Delta_z}, & \frac{\partial \widetilde{\theta}}{\partial z} < 0 , \\
    0.76\, \sqrt{k}\, \frac{g}{\theta_\circ}\, \left( \frac{\partial \widetilde{\theta}}{\partial z}\right)^{-1/2}, &  \frac{\partial \widetilde{\theta}}{\partial z} > 0 .\\
  \end{cases}
\end{align}

\subsubsection{\label{sec:wallmodelBC}Lower wall model BC}
The sub-filter scale stresses are specified at the terrain boundary following
Moeng~\cite{Moeng1984} as
\begin{align}
 (\tau_{xz})_0 &= \average{\tau_{xz}}_0 \frac{V_1 \average{\widetilde{u}_1} + \average{V_1} (\widetilde{u}_1 - \average{\widetilde{u}_1} ) }{ \average{V_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } , \\
 (\tau_{yz})_0 &= \average{\tau_{yz}}_0 \frac{V_1 \average{\widetilde{v}_1} + \average{V_1} (\widetilde{v}_1 - \average{\widetilde{v}_1} ) }{ \average{V_1} \sqrt{\average{\widetilde{u}_1}^2 + \average{\widetilde{v}_1}^2 } } , \\
  (\tau_{\theta z})_0 &= \average{\tau_{\theta z}}_0 \frac{V_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{V_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{V_1} ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) },
\end{align}
where all quantities with the subscript 1 are values at the first vertical grid
level, $V_1$ is the resolved wind speed at the first grid level and
$\average{\bullet}$ represents the horizontal plane average at that height. The mean
sub-filter stresses are computed using the Monin-Obukhov similarity
law~\cite{Etling1996}. The formulation for the potential temperature flux could
potentially experience a singularity when the mean resolved potential
temperature at the first grid level $\average{\widetilde{\theta}_1}$ crosses the
reference temperature $\theta_{\circ}$. Hence the model for the sub-filter
potential temperature flux is reformulated using the Monin-Obukhov similarity
law as
\begin{equation}
(\tau_{\theta z})_0 = \left ( \frac{u_{\tau} \kappa}{ \phi_h} \right ) \frac{V_1 ( \average{\widetilde{\theta}_1} - \theta_{\circ} ) + \average{V_1} (\widetilde{\theta}_1 - \average{\widetilde{\theta}_1} ) }{ \average{V_1} },
\end{equation}
where $\phi_h$ is the non-dimensional potential temperature gradient from
Monin-Obukhov similarity law~\cite{Etling1996}.

>>>>

\subsubsection{Governing equations}
Both computational fluid dynamics (CFD) codes solve the incompressible form of
the Navier-Stokes equations with appropriate models for turbulence
closure. Equation~\ref{eqn:ns-les} shows the continuity and momentum equations
with all the terms necessary to model the atmospheric boundary layer
problem. LES filtered variables are denoted using the operation ${\left(\bullet \right)}$.
Term $\mathbf{I}$ represents the pressure gradient forces as a
deviation from hydrostatic and horizontal mean gradients, term $\mathbf{II}$
represents the contribution from viscous and sub-filter scale stresses, term
$\mathbf{III}$ represents the contribution from Coriolis forces due to earth's
rotation, term $\mathbf{IV}$ represents the effects of buoyancy, and term
$\mathbf{V}$ represents source terms necessary to drive the flow to a desired
horizontal mean velocity at prescribed heights.

\begin{align}
  \frac{\partial {u_j}}{\partial x_j} & = 0 \nonumber\\
  \frac{\partial}{\partial t} \left(\rho\, {u}_i\right) +
  \frac{\partial}{\partial x_j} \left( \rho\, {u}_i {u}_j \right) &=
  - \underbrace{\frac{\partial p'}{\partial x_j} \delta_{ij}}_\mathbf{I}
  - \underbrace{\frac{\partial \tau_{ij}}{\partial x_j}}_\mathbf{II}
  - \underbrace{2\rho\,\epsilon_{ijk}\,\Omega_j{u_k}}_\mathbf{III}
  + \underbrace{\left(\rho - \rho_\circ \right) g_i}_\mathbf{IV}
  + \underbrace{S^{u}_{i}}_\mathbf{V} \label{eqn:ns-les}
\end{align}
For the simulations presented in this paper, the buoyancy term ($\mathbf{IV}$)
is approximated using the Bousinessq model. The Bousinessq model ignores the
effect of density gradients on inertia while retaining its effects on buoyancy.
The density fluctuations governing the buoyancy effects are approximated as

\begin{align}
  \frac{\rho}{\rho_\circ} \approx 1 - \beta \left( \theta - \theta_\circ \right)
\end{align}
This leads to a forcing term that depends on potential temperature ($\theta$),
the reference density $\rho_\circ = \rho$, and thermal expansion
coefficient $\beta \approx 1 / \theta_\circ$ as
\begin{align}
  \left(\rho - \rho_\circ \right) g_i = -\rho_\circ\, \beta\, g_i \left( \theta - \theta_\circ \right)
\end{align}

For ABL problems, the conservation of energy equation is usually written in
terms of potential temperature as shown in Eq.~\ref{eqn:pot-temp-les}.
\begin{align}
  \frac{\partial}{\partial t} \left(\rho\, {\theta}\right) +
  \frac{\partial}{\partial x_j} \left(\rho\, {u}_j {\theta} \right) = - \frac{\partial}{\partial x_j} \hat{q}_j \label{eqn:pot-temp-les}
\end{align}
Under the assumption of ideal gas conditions and constant specific heat capacity
$c_p$, the gradients in potential temperature are proportional to the gradients
in absolute temperature, i.e.,
\begin{align}
   \left[ \frac{\partial T}{\partial t}, \frac{\partial T}{\partial x}, \frac{\partial T}{\partial y} \right] =
   \left( \frac{\bar{p}}{p_\circ} \right)^{\left(\frac{R}{c_p}\right)} \left[ \frac{\partial \theta}{\partial t}, \frac{\partial \theta}{\partial x}, \frac{\partial \theta}{\partial y} \right]
\end{align}
Furthermore, ignoring the pressure and viscous work terms, for incompressible
flows, it can be shown that solving the enthalpy equation is equivalent to
solving the potential temperature equation. The enthalpy equation solved in
Nalu-Wind for wind energy problems is shown below
\begin{align}
  \frac{\partial}{\partial t} \left(\rho\, {h}\right) +
  \frac{\partial}{\partial x_j} \left(\rho\, {u}_j {h} \right) = - \frac{\partial}{\partial x_j} q_j . \label{eqn:enthalpy-les}
\end{align}
It is noted here that the terms $\hat{q}_j$ (Eq.~\ref{eqn:pot-temp-les}) and
$q_j$ (Eq.~\ref{eqn:enthalpy-les}) are not equivalent and must be scaled
appropriately. Users can still provide the appropriate initial and boundary
conditions in terms of potential temperature field. Under these assumptions and
conditions, the resulting solution can then be interpreted as the variation of
potential temperature field in the computational domain.

@@

\begin{align}
  \frac{\partial}{\partial t}\left(\rho k \right) + \frac{\partial}{\partial x_j} \rho k {u_j} &= \frac{\partial}{\partial x_j} \left (2 \nu_t \frac{\partial k}{\partial x_j} \right) - 2 \nu_t {S_{ij}} {S_{ij}} + \frac{g}{\theta_\circ} \tau_{\theta_i} - C_{\epsilon} \frac{k^{3/2}}{l} \label{eqn:ksgs-les}
\end{align}
where
\begin{align*}
  C_\epsilon &= 0.93 & \nu_t &= C_\epsilon l \sqrt{k} \\
  \tau_{\theta i} &= - 2\ \frac{\nu_t}{\mathrm{Pr}_t}\ \frac{\partial {\theta}}{\partial x_i} & \mathrm{Pr}_t &= \left( 1 + \frac{2l}{\Delta} \right)^{-1}
\end{align*}
The choice of the length scale ($l$) depends on whether the stratification is positive
or negative, based on the recommendation by Moeng~\cite{Moeng1984}, and is given
by
\begin{align}
  l =
  \begin{cases}
    \Delta = \sqrt[3]{\Delta_x \Delta_y \Delta_z}, & \frac{\partial {\theta}}{\partial z} < 0 , \\
    0.76\, \sqrt{k}\, \frac{g}{\theta_\circ}\, \left( \frac{\partial {\theta}}{\partial z}\right)^{-1/2}, &  \frac{\partial {\theta}}{\partial z} > 0 .\\
  \end{cases}
\end{align}

\subsubsection{\label{sec:wallmodelBC}Lower wall model BC}
The sub-filter scale stresses are specified at the terrain boundary following
Moeng~\cite{Moeng1984} as
\begin{align}
 (\tau_{xz})_0 &= \average{\tau_{xz}}_0 \frac{V_1 \average{{u}_1} + \average{V_1} ({u}_1 - \average{{u}_1} ) }{ \average{V_1} \sqrt{\average{{u}_1}^2 + \average{{v}_1}^2 } } , \\
 (\tau_{yz})_0 &= \average{\tau_{yz}}_0 \frac{V_1 \average{{v}_1} + \average{V_1} ({v}_1 - \average{{v}_1} ) }{ \average{V_1} \sqrt{\average{{u}_1}^2 + \average{{v}_1}^2 } } , \\
  (\tau_{\theta z})_0 &= \average{\tau_{\theta z}}_0 \frac{V_1 ( \average{{\theta}_1} - \theta_{\circ} ) + \average{V_1} ({\theta}_1 - \average{{\theta}_1} ) }{ \average{V_1} ( \average{{\theta}_1} - \theta_{\circ} ) },
\end{align}
where all quantities with the subscript 1 are values at the first vertical grid
level, $V_1$ is the resolved wind speed at the first grid level and
$\average{\bullet}$ represents the horizontal plane average at that height. The mean
sub-filter stresses are computed using the Monin-Obukhov similarity
law~\cite{Etling1996}. The formulation for the potential temperature flux could
potentially experience a singularity when the mean resolved potential
temperature at the first grid level $\average{{\theta}_1}$ crosses the
reference temperature $\theta_{\circ}$. Hence the model for the sub-filter
potential temperature flux is reformulated using the Monin-Obukhov similarity
law as
\begin{equation}
(\tau_{\theta z})_0 = \left ( \frac{u_{\tau} \kappa}{ \phi_h} \right ) \frac{V_1 ( \average{{\theta}_1} - \theta_{\circ} ) + \average{V_1} ({\theta}_1 - \average{{\theta}_1} ) }{ \average{V_1} },
\end{equation}
where $\phi_h$ is the non-dimensional potential temperature gradient from
Monin-Obukhov similarity law~\cite{Etling1996}.
